Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-06-14 GWLDCI.tex
addressed Sian's comments on 3.25

By the third point of Lemma \ref{maxexactcomplementrank}, the complementary flat, $F(P)$ has full rank: $\rk(F(P)) = |P|$. Let $B \subset F(P)$ be a maximal independent subset of $F(P)$: $\rk (B) = |P|$. Since $F(P)$ is a flat, adding any element of $V(R)$ to $B$ increases the rank. Therefore, any basis of $W$ can be written as $B \cup U$ for some $B$ and $U$ of the form indicated. However, since $F(P)$ is common to both $W$ and $W'$, and $V(R) = V(R')$, any any basis of $W'$ can also be written $B \cup U$. Thus both matroids have the same bases sets, proving that they are the same.

For the converse, assume that the matroids associated to $W$ and $W'$ are the same: $M(W) = M(W')= M$. Let $\mathcal{R} = \{(P_i, V(P_i)\}_{i=1}^m$ and $\mathcal{R}' = \{(P'_i, V(P'_i)\}_{i=1}^l$ be the sets of maximal exact subdiagrams of $W$ and $W'$. Recall that these sets $\{P_i\}$ and $\{P_i'\}$ partition the propagator sets $\cP$ and $\cP'$ of $W$ and $W'$ respectivley. Let $\mathcal{F}$ and $\mathcal{F}'$ the respective complementary vertices: \bas \mathcal{F} = \{F_i = V(P_i)^c | (P_i, V(P_i)) \in \mathcal{R}\} \quad  \mathcal{F}' = \{F_i' = V(P'_i)^c | (P'_i, V(P'_i)) \in \mathcal{R}\} \;.\eas By Theorem \ref{maxexactcomplementrank}, every $F_i$ and $F_i'$ is a cyclic flat of full rank in $M$. We claim that $\mathcal{F} = \mathcal{F'}$.

If not, there would be a cyclic flat $F_j$ of $M$. Without loss of generality, write $F_j \in \mathcal{F} \setminus \mathcal{F'}$.  We may write $F_j = V(P_j)^c$, with $(P_j, V(P_j))$ a maximal exact subdiagram in $W$. By Lemma \ref{lem facts about WLD matroids} (5), we may write $F_j= F(Q)$ as a propagator flat. By Lemma \ref{contractsubdiaglem}, $M/F(Q) = F/F_j = M(Q^c, V(Q^c))$, which is uniform, by construction of $F_j$. Theorem \ref{exactuniformthm} $(Q^c, V(Q^c))$ is exact. Since this exact subdiagram is not in $\mathcal{R}'$ by construction of $F_j$, it is not a maximal exact subdiagram of $W'$. That is, there is a maximal exact subdiagram containing it: $(Q^c, V(Q^c)) \subset (P_j', V(P_j'))$. Suppose $|P_j' \setminus Q^c| = m$. By counting vertices of the two nested sets $F_j' \subset F_j$ we see that  \bas |F_j| = n -|Q^c| -3  \textrm{ while } |F_j'| = n -|P_j'| -3 = n -|Q^c| -m -3\;.\eas That is, $F_j'$ has $m$ more vertices than $F_j$. Furthermore, by construction $\rk(F_j) + m = \rk (F_j')$. This violates the assumption that both $F_j$ and $F_j'$ are cyclic flats.

Therefore, we see that $\mathcal{F} = \mathcal{F'}$, which implies that the maximally exact subdiagrams of $W$ and $W'$, are supported on the same vertex sets. All propagators of $W$ and $W'$ are accounted for in these sets. There there exists a transitive chain of equivalent diagrams that put $W$ and $W'$ in the same equivalence class.
\end{proof}

>>>>

By the third point of Lemma \ref{maxexactcomplementrank}, the complementary flat, $F(P)$ has rank equal to the number of propagators: $\rk(F(P)) = |P|$. Let $B \subset F(P)$ be a maximal independent subset of $F(P)$: $\rk (B) = |P|$. Since $F(P)$ is a flat, adding any element of $V(R)$ to $B$ increases the rank. Therefore, any basis of $W$ can be written as $B \cup U$ for some $B$ and $U$ of the form indicated. However, since $F(P)$ is common to both $W$ and $W'$, and $V(R) = V(R')$, any any basis of $W'$ can also be written $B \cup U$. Thus both matroids have the same bases sets, proving that they are the same.

For the converse, assume that the matroids associated to $W$ and $W'$ are the same: $M(W) = M(W')= M$. Let $\mathcal{R} = \{(P_i, V(P_i)\}_{i=1}^m$ and $\mathcal{R}' = \{(P'_i, V(P'_i)\}_{i=1}^l$ be the sets of maximal exact subdiagrams of $W$ and $W'$. Recall that these sets $\{P_i\}$ and $\{P_i'\}$ partition the propagator sets $\cP$ and $\cP'$ of $W$ and $W'$ respectivley. Let $\mathcal{F}$ and $\mathcal{F}'$ the respective complementary vertices: \bas \mathcal{F} = \{F_i = V(P_i)^c | (P_i, V(P_i)) \in \mathcal{R}\} \quad  \mathcal{F}' = \{F_i' = V(P'_i)^c | (P'_i, V(P'_i)) \in \mathcal{R}\} \;.\eas By Theorem \ref{maxexactcomplementrank}, every $F_i$ and $F_i'$ is a cyclic flat in $M$. They are also propagator flats in $W$ and $W'$ of rank equal to the number of propagators defining it. We claim that $\mathcal{F} = \mathcal{F'}$.

If not, there would be a cyclic flat $F_j$ of $M$. Without loss of generality, write $F_j \in \mathcal{F} \setminus \mathcal{F}'$.  We may write $F_j = V(P_j)^c$, with $(P_j, V(P_j))$ a maximal exact subdiagram in $W$. By Lemma \ref{lem facts about WLD matroids} (5), we may write $F_j= F(Q)$ as a propagator flat. Therefore, Lemma \ref{maxexactcomplementrank}, $\rk(F_j) = |Q|$. By Lemma \ref{contractsubdiaglem}, $M/F(Q) = F/F_j = M(Q^c, V(Q^c))$, which is uniform, by construction of $F_j$. Theorem \ref{exactuniformthm} $(Q^c, V(Q^c))$ is exact. Since this exact subdiagram is not in $\mathcal{R}'$ by construction of $F_j$, it is not a maximal exact subdiagram of $W'$. That is, there is a maximal exact subdiagram containing it: $(Q^c, V(Q^c)) \subsetneq (P_j', V(P_j'))$. Thus there is a $F_j' \in \mathcal{R}'$ such that $F_j' \subsetneq F_j$. We show that $F_j$ is not a cyclic flat.

Suppose the exact diagram  $(P_j', V(P_j'))$ has $m$ more propagators than $(Q^c, V(Q^c))$: $|P_j' \setminus Q^c| = m$. Since $(P_j', V(P_j'))$ defines a uniform matroid, any subset of vertices numbering fewer than the number of propagators in $P_j'$ forms an independent flat of $M(P_j', V(P_j'))$, and thus an independent set of $M$. \sanote{is it worth citing a result that says for $W \subset W'$, if $V$ independent in $M(W)$ then also independent in $M(W')$?} Next, compare the number of vertices in $F_j'$ and $F_j$ to see that $F_j'$ has $m$ more vertices than $F_j$:  \bas |F_j| = n -|Q^c| -3  \textrm{ while } |F_j'| = n -|P_j'| -3 = n -|Q^c| -m -3\;.\eas So $|F_j \setminus F_j' | = m$ and $F_j \setminus F_j' = V(P_j')\setminus V(Q^c)$.  Thus $F_j \setminus F_j'$ is an independent set in $M$. Therefore $F_j$ is the union of a cyclic flat, $F_j'$, and an independent set, $F_j \setminus F_j'$. Therefore, $F_j$ is not a cyclic flat. Therefore, $F_j \not \in \mathcal{F} \setminus \mathcal{F}'$. Therefore, $\mathcal{F} = \mathcal{F}'$, as claimed.

Since $\mathcal{F} = \mathcal{F'}$, the maximally exact subdiagrams of $W$ and $W'$, are supported on the same vertex sets. All propagators of $W$ and $W'$ are accounted for in these sets. There there exists a transitive chain of equivalent diagrams that put $W$ and $W'$ in the same equivalence class.
\end{proof}
